---
title: "Clustered Standard Errors With (Too) Few Clusters"
date: '2026-09-07'
permalink: /posts/2026/09/few-clusters/
excerpt: "A simulation of 95% confidence interval coverage under regular, cluster-robust, CR2/Satterthwaite, and wild cluster bootstrap inference, as the number of clusters shrinks from 50 to 2."
tags:
  - econometrics
  - clustered standard errors
  - bootstrap
  - simulation
# --- render config (stripped from the post by render_post.R) ---
output:
  md_document:
    variant: gfm
    preserve_yaml: TRUE
knit: (function(inputFile, encoding) { source("render_post.R"); render_post(inputFile, encoding) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      dev = "ragg_png", dpi = 110, fig.width = 10, fig.height = 6)
```

Back in April 2021 I posted a [Twitter thread](https://twitter.com/gburtch/status/1378520203689082886) with a small simulation of what happens to cluster-robust standard errors when you only have a handful of clusters. It got passed around more than I expected, and Alexander Fischer (author of the excellent `fwildclusterboot` package) wrote a [nice follow-up](https://s3alfisc.github.io/blog/post/2022-01-29-cluster-robust-inference-when-the-number-of-clusters-is-small-a-horse-race/) that added Satterthwaite-corrected standard errors to the comparison. The thread never made it onto this site, and the original code has aged (it used `multiwayvcov`, which has since left CRAN). So here is an updated version, with the small-sample corrections folded in.

## The problem

We cluster standard errors because observations within a group (a state, a firm, a classroom, a subreddit) share unobserved shocks, so regression errors are correlated within the group. Ignore that and you understate your standard errors, sometimes badly ([Moulton, 1986](https://doi.org/10.1016/0304-4076(86)90021-7)). The standard fix is the cluster-robust "sandwich" variance estimator of [Liang and Zeger (1986)](https://doi.org/10.1093/biomet/73.1.13) (what I will call vanilla cluster SEs), which is consistent as the number of clusters, $G$, goes to infinity.

The catch is that $G$ is frequently not large. Papers cluster on 50 states, on a dozen regions, or on the six sites of a field experiment. Two questions follow: how badly do clustered standard errors perform when $G$ is small, and what can we do about it? The answer to the second question, at least for the linear model, is well developed: use a better small-sample adjustment ([Bell and McCaffrey, 2002](https://www150.statcan.gc.ca/n1/en/catalogue/12-001-X20020026291); [Imbens and Kolesár, 2016](https://doi.org/10.1162/REST_a_00552); [Pustejovsky and Tipton, 2018](https://doi.org/10.1080/07350015.2016.1247004)) or use the wild cluster bootstrap ([Cameron, Gelbach and Miller, 2008](https://doi.org/10.1162/rest.90.3.414); [Roodman et al., 2019](https://doi.org/10.1177/1536867X19830877)). [MacKinnon, Nielsen and Webb (2023)](https://doi.org/10.1016/j.jeconom.2022.04.001) is a good current guide to all of this. The point of this post is just to *see* it.

## The data generating process

I simulate a bivariate regression $y = 0.1 + 0.5x + e$ with $n = 1000$ observations split evenly across $G$ clusters. Both the regressor and the error have a cluster-level component:

$$x_{ig} = x^{(i)}_{ig} + x^{(g)}_{g}, \qquad e_{ig} = e^{(i)}_{ig} + e^{(g)}_{g},$$

where the individual-level pieces are independent normal draws and the cluster-level pieces are shared by everyone in cluster $g$. I set the intra-cluster correlation of the error to $\rho = 0.7$ by giving $e^{(g)}$ variance $\rho$ and $e^{(i)}$ variance $1-\rho$. The cluster component of $x$ matters too: the Moulton inflation factor for the standard error scales with the *product* of the intra-cluster correlations of $x$ and $e$ (and with cluster size), so a clustered error term alone does not do much damage if $x$ varies freely within clusters. Note that $x$ and $e$ are independent, so there is no endogeneity here.

```{r dgp}
suppressPackageStartupMessages({
  library(dplyr); library(ggplot2); library(patchwork)
  library(sandwich); library(clubSandwich); library(fwildclusterboot); library(future.apply)
})

gen_cluster <- function(n = 1000, G = 50, rho = 0.7, beta = c(0.1, 0.5)) {
  n <- ceiling(n / G) * G                      # balanced clusters
  ind <- mvtnorm::rmvnorm(n, sigma = diag(c(1, 1 - rho)))   # individual parts of (x, e)
  cl  <- mvtnorm::rmvnorm(G, sigma = diag(c(1, rho)))       # cluster parts of (x, e)
  cluster <- rep(seq_len(G), each = n / G)
  x <- ind[, 1] + cl[cluster, 1]
  e <- ind[, 2] + cl[cluster, 2]
  data.frame(x = x, y = beta[1] + beta[2] * x + e, cluster = cluster)
}
```

## 4 approaches to constructing cluster SEs

For every simulated dataset I fit `lm(y ~ x)` once and then construct standard errors using each approach:

1. **Regular OLS.** The textbook standard error, which assumes independent errors.
2. **CR1, $t(n-k)$.** Vanilla cluster SEs: the cluster-robust sandwich with Stata's $\frac{G}{G-1}\frac{n-1}{n-k}$ finite-sample factor, paired with the same critical value OLS would use. This is what R's `sandwich::vcovCL()` followed by `lmtest::coeftest()` gives you by default.
3. **CR1, $t(G-1)$.** Identical variance estimate, but with critical values from a $t$ distribution with $G-1$ degrees of freedom. This is the default in Stata's `vce(cluster)` and in R's `fixest`.
4. **CR2 + Satterthwaite.** The bias-reduced CR2 sandwich of Bell and McCaffrey (2002), with Satterthwaite degrees of freedom, via `clubSandwich`.
5. **Wild cluster bootstrap.** 999 bootstrap draws imposing the null, via `fwildclusterboot`. I use Rademacher weights, switching to Webb's six-point weights when $G \le 12$, as Roodman et al. (2019) recommend. The confidence interval is obtained by test inversion.

```{r estimators}
one_sim <- function(G, n = 1000, rho = 0.7, B = 999, seed = 1) {
  set.seed(seed); dqrng::dqset.seed(seed)
  d <- gen_cluster(n = n, G = G, rho = rho)
  fit <- lm(y ~ x, data = d)
  b <- unname(coef(fit)["x"]); n_obs <- nrow(d)

  ci_reg <- confint(fit)["x", ]

  V1  <- sandwich::vcovCL(fit, cluster = ~cluster, type = "HC1")
  se1 <- sqrt(V1["x", "x"])
  ci_cr1_n <- b + c(-1, 1) * qt(0.975, n_obs - 2) * se1
  ci_cr1_g <- b + c(-1, 1) * qt(0.975, G - 1) * se1

  ci_cr2 <- tryCatch({
    ct <- clubSandwich::conf_int(fit, vcov = "CR2", cluster = d$cluster,
                                 test = "Satterthwaite", coefs = "x")
    c(ct$CI_L, ct$CI_U)
  }, error = function(e) c(NA, NA))

  ci_wild <- tryCatch({
    bt <- suppressWarnings(suppressMessages(fwildclusterboot::boottest(
      fit, param = "x", clustid = "cluster", B = B,
      type = if (G <= 12) "webb" else "rademacher")))
    bt$conf_int
  }, error = function(e) c(NA, NA))

  data.frame(G = G, b = b,
             method = c("regular", "cr1_t_n", "cr1_t_g", "cr2_satt", "wild"),
             lo = c(ci_reg[1], ci_cr1_n[1], ci_cr1_g[1], ci_cr2[1], ci_wild[1]),
             hi = c(ci_reg[2], ci_cr1_n[2], ci_cr1_g[2], ci_cr2[2], ci_wild[2]))
}
```

## Running it

I run 1000 replications for every cluster count from 2 to 50.

```{r simulate, cache=TRUE}
run_sims <- function(G_grid, n_sims = 1000, n = 1000, rho = 0.7, B = 999,
                     workers = max(1, parallel::detectCores() - 4)) {
  grid <- expand.grid(G = G_grid, sim = seq_len(n_sims))
  plan(multisession, workers = workers); on.exit(plan(sequential), add = TRUE)
  res <- future_lapply(seq_len(nrow(grid)), function(i) {
    r <- one_sim(G = grid$G[i], n = n, rho = rho, B = B, seed = 1e6 + i)
    r$sim <- grid$sim[i]
    r
  }, future.seed = TRUE)
  bind_rows(res)
}

sims <- run_sims(G_grid = 2:50) %>%
  mutate(covered = lo <= 0.5 & hi >= 0.5,
         method = factor(method, levels = c("regular", "cr1_t_n", "cr1_t_g", "cr2_satt", "wild"),
                         labels = c("Regular OLS", "CR1, t(n-k): R sandwich + coeftest default", "CR1, t(G-1): Stata / R fixest default",
                                    "Clustered CR2 + Satterthwaite", "Wild cluster bootstrap")))

coverage <- sims %>%
  group_by(G, method) %>%
  summarise(coverage = mean(covered, na.rm = TRUE),      # coverage among intervals that were produced
            failed = mean(is.na(covered)),               # share of replications with no interval (bootstrap only)
            med_width = median(hi - lo, na.rm = TRUE), .groups = "drop")
```

```{r helpers, include=FALSE}
cv <- function(m, g) sprintf("%.1f%%", 100 * coverage$coverage[coverage$G == g & coverage$method == m])
fl <- function(m, g) sprintf("%.0f%%", 100 * coverage$failed[coverage$G == g & coverage$method == m])
wd <- function(m, g) sprintf("%.2f", coverage$med_width[coverage$G == g & coverage$method == m])
theme_post <- function() {
  theme_classic(base_family = "Economica", base_size = 16) +
    theme(plot.title = element_text(size = 15), legend.position = "bottom")
}
plot_cis <- function(d, title) {
  ggplot(filter(d, !is.na(covered)), aes(x = sim, y = b, ymin = lo, ymax = hi, color = covered)) +
    geom_errorbar(color = "gray80", alpha = 0.6, width = 0) +
    geom_point(size = 1.1) +
    geom_hline(yintercept = 0.5, linetype = "dashed", color = "chartreuse4") +
    scale_color_manual(values = c(`TRUE` = "blue", `FALSE` = "red"),
                       labels = c(`TRUE` = "Yes", `FALSE` = "No"), name = "95% CI covers the true beta?") +
    labs(x = "Simulation", y = "Estimate (95% CI)", title = title) +
    theme_post()
}
```

## Vanilla clustering with a sufficiently large number of clusters

Start with $G = 50$. The top panel uses regular OLS standard errors; the bottom uses vanilla cluster SEs. Each point in the plot reflects one point estimate from a single simulated dataset. Red points are those where the 95% CI does not include the true value of 0.5 I specified in my data-generating process.

```{r g50, echo=FALSE, fig.height=9}
d50 <- filter(sims, G == 50)
p1 <- plot_cis(filter(d50, method == "Regular OLS"),
               sprintf("Regular OLS standard errors (G = 50, rho = 0.7): coverage %s", cv("Regular OLS", 50)))
p2 <- plot_cis(filter(d50, method == "CR1, t(n-k): R sandwich + coeftest default"),
               sprintf("Clustered (CR1) standard errors (G = 50, rho = 0.7): coverage %s", cv("CR1, t(n-k): R sandwich + coeftest default", 50)))
(p1 / p2) + plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

With regular standard errors the "95%" interval covers the truth `r cv("Regular OLS", 50)` of the time. The estimates themselves are unbiased but the 95% CI is too narrow. Clustering fixes most of the problem: coverage rises to `r cv("CR1, t(n-k): R sandwich + coeftest default", 50)`.

## Six clusters

What happens when we cut $G$ from 50 to 6? Each cluster now has about 167 observations. The top panel uses vanilla cluster SEs; the bottom the wild cluster bootstrap.

```{r g6, echo=FALSE, fig.height=9}
d6 <- filter(sims, G == 6)
p3 <- plot_cis(filter(d6, method == "CR1, t(n-k): R sandwich + coeftest default"),
               sprintf("Clustered (CR1) standard errors (G = 6, rho = 0.7): coverage %s", cv("CR1, t(n-k): R sandwich + coeftest default", 6)))
p4 <- plot_cis(filter(d6, method == "Wild cluster bootstrap"),
               sprintf("Wild cluster bootstrap (G = 6, rho = 0.7): coverage %s", cv("Wild cluster bootstrap", 6)))
(p3 / p4) + plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

With six clusters vanilla cluster SEs cover the truth only `r cv("CR1, t(n-k): R sandwich + coeftest default", 6)` of the time; regular standard errors are hopeless at `r cv("Regular OLS", 6)`. The wild cluster bootstrap gets to `r cv("Wild cluster bootstrap", 6)`. Its intervals are somewhat wider and more variable in width.

## Coverage as a function of the number of clusters

Here is the full picture, for every $G$ from 2 to 50 and all five procedures.

```{r coverage_plot, echo=FALSE, fig.height=7}
ggplot(coverage, aes(x = G, y = 100 * coverage, color = method)) +
  geom_hline(yintercept = 95, linetype = "dashed", color = "red") +
  geom_line(linewidth = 1.3, alpha = 0.85) +
  scale_x_continuous(breaks = c(2, 5, 10, 20, 30, 40, 50)) +
  scale_y_continuous(breaks = seq(0, 100, 10), limits = c(0, 100)) +
  scale_color_manual(values = c("darkorange", "steelblue", "navy", "purple", "forestgreen"), name = NULL) +
  guides(color = guide_legend(nrow = 3)) +
  labs(x = "Number of clusters (1000 simulations each; n = 1000)",
       y = "Coverage of the true slope by the 95% CI (%)",
       title = "Confidence interval coverage by number of clusters (rho = 0.7)") +
  theme_post()
```

```{r coverage_table, echo=FALSE}
coverage %>%
  filter(G %in% c(2, 3, 4, 5, 6, 8, 10, 15, 20, 30, 50)) %>%
  mutate(coverage = sprintf("%.1f", 100 * coverage)) %>%
  select(G, method, coverage) %>%
  tidyr::pivot_wider(names_from = method, values_from = coverage) %>%
  knitr::kable(caption = "Coverage (%) of nominal 95% confidence intervals.")
```

A few things stand out.

1. **Regular standard errors get *worse* as clusters get fewer.** With $n$ fixed, fewer clusters means bigger clusters, and the Moulton factor grows with cluster size. At $G = 2$ coverage is `r cv("Regular OLS", 2)`; even at $G = 50$ it is only `r cv("Regular OLS", 50)`.
2. **Vanilla cluster SEs lack coverage below 30 or so clusters, and badly below 10.** At $G = 10$ it covers `r cv("CR1, t(n-k): R sandwich + coeftest default", 10)`; at $G = 5$, `r cv("CR1, t(n-k): R sandwich + coeftest default", 5)`.
3. **Using $t(G-1)$ critical values is free and helps.** Some software now uses this by default (Stata and `fixest` do).
4. **CR2 with Satterthwaite degrees of freedom and the wild cluster bootstrap are both effective down to about five clusters.** At $G = 6$ they cover `r cv("Clustered CR2 + Satterthwaite", 6)` and `r cv("Wild cluster bootstrap", 6)` respectively.
5. **Below five clusters, all bets are off.** With two clusters the wild bootstrap fails to return an interval at all in `r fl("Wild cluster bootstrap", 2)` of replications (Webb weights give only $6^G = 36$ distinct draws) and covers `r cv("Wild cluster bootstrap", 2)` when it does. CR2 with Satterthwaite goes the other way: the degrees of freedom collapse toward one, the median interval is `r wd("Clustered CR2 + Satterthwaite", 2)` wide (the slope is 0.5), and coverage is `r cv("Clustered CR2 + Satterthwaite", 2)`.

## Takeaways

If you cluster, know how many clusters you have. Above 40 or 50, vanilla cluster SEs are fine. Between roughly 10 and 40, at minimum use $t(G-1)$ critical values, and preferably CR2 or the wild cluster bootstrap. Below 10, use CR2 or the bootstrap and report the interval, not just the stars. And below five, be honest with yourself and your readers that the data cannot tell you much about sampling variability at the cluster level.

The original 2021 code is archived in [this repository](https://github.com/gburtch/simulating_cluster_SEs); the version above supersedes it.
